---
title: "Summary Values"
author: "Jasen"
date: "2024-09-05"
output: pdf_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(gridExtra)
library(geepack)
library(scales)
library(reshape2)
library(lsmeans)
dir <- getwd()
codes_dir <- paste(substr(dir,1, nchar(dir)-4), "Code/00_Codes.R", sep = '')
source(codes_dir)
```

# load the data

```{r}
dir <- getwd()
codes_dir <- paste(substr(dir,1, nchar(dir)-4), "Data/long_temporal_data_V012522.RData", sep = '')
load(codes_dir)
```

# How many children

```{r}
table(df2_long$Index)
```

# How many children in each time-period

```{r}
wow <- table(df2_long$nid, df2_long$Index)

covariates <- c('age',  'race','gender', 'Index', 'days_after_mdh', 'zha', 'i_g4salary_scaled', 'i_d9educatpar', 'currentnoflorag')

df3_long <- df2_long[!is.na(df2_long$age),]
df3_long <- df2_long[!is.na(df2_long$dist_house_plant),]
df4_long <- df3_long %>% filter(!if_any(covariates, is.na))

table(df3_long$Index)
table(df4_long$Index)

round(100 * table(df3_long$Index) / sum(df2_long$Index == 1), 2)


# how many missing values for the outcomes
NEPSY <- colnames(df4_long)[47:63]
df4_long_NEPSY <- df4_long %>% select(NEPSY) 

df4_long <- df4_long[!apply(df4_long_NEPSY, 1, function(x) all(is.na(x))), ]

table(df4_long$Index)

length(unique(df4_long$nid))

missing_NID <- setdiff(unique(df2_long$nid), unique(df4_long$nid))

df5_long <- df2_long %>% filter(nid %in% missing_NID)
```
# how many in 2008 came back in spring 2016

```{r}
df3_2008 <- df3_long %>% filter(Index == 1)
ID_2008 <- df3_2008$nid
df3_2008_v2 <- df3_long %>% filter(nid %in% ID_2008)

table(df3_2008_v2$Index)

# how many new recruits not in 2008

ID_not_2008 <- setdiff(unique(df3_long$nid), ID_2008)

df3_newcomers <- df3_long %>% filter(nid %in% ID_not_2008)

table(df3_newcomers$Index)

table(df3_2008_v2$Index) + table(df3_newcomers$Index)
```


# venn diagram of membership

```{r}
# 629 unique children
length(unique(df3_long$nid))

wow <- tapply(df3_long$Index, factor(df3_long$nid), table)

values <- data.frame()

for(i in 1:length(wow)){
  values <- rbind(values, unname(wow[[i]]))
}

colnames(values) <- c('Summer 2008', 'Spring 2016', 'Summer 2016')

row_counts <- values %>%
  group_by(across(everything())) %>%
  summarise(Frequency = n(), .groups = 'drop')

print(row_counts)

sum(row_counts$Frequency)
```

# how many ppl in an actual aac model

- df2_long = 1887 = 629 * 3

```{r}
covariates <- c('dist_house_plant_log',
                                  'age',
                                  'race',
                                  'gender',
                                  'days_after_mdh',
                                  'zha',
                                  'i_g4salary_scaled',
                                  'i_d9educatpar',
                                  'currentnoflorag')

df_model <- df2_long %>% select(c('nid',
                                  'Index',
                                  'dist_house_plant_log',
                                  'age',
                                  'race',
                                  'gender',
                                  'days_after_mdh',
                                  'zha',
                                  'i_g4salary_scaled',
                                  'i_d9educatpar',
                                  'currentnoflorag',
                                  'aac'))

table(df_model$Index)
length(unique(df_model$nid))

df_model2 <- df_model %>% filter(!is.na(dist_house_plant_log))
table(df_model2$Index)
length(unique(df_model2$nid))

df_model3 <- df_model2 %>% filter(!is.na(aac))
table(df_model3$Index)
length(unique(df_model3$nid))

# any missing covariates? Yes
apply(df_model3, 2, function(x){sum(is.na(x))})

df_model4 <- df_model3 %>% na.omit()

table(df_model4$Index)

length(unique(df_model4$nid))
```

# distribution of 2008 cohabitation

```{r}
df_summer_08 <- df2_long %>% filter(Index == 1) %>% 
  filter(!is.na(age))

table(df_summer_08$currentnoflorag)

round(100 * table(df_summer_08$currentnoflorag) / nrow(df_summer_08), 2)
```

# TABLE 2 CALCULATIONS (9/19/2025)

```{r}

df3_long <- df4_long

cont_calc <- function(var_name){
  r1 <- round(tapply(df3_long[[var_name]], df3_long$Index, function(x){mean(x, na.rm = T)}), 1)
  
  r2 <- round(tapply(df3_long[[var_name]], df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
  
  result <- matrix(paste0(r1, ' (', r2, ')'), nrow = 1)
  rownames(result) <- var_name
  
  return(result)
}



cat_calc <- function(var_name){
  counts <- table(df3_long[[var_name]], df3_long$Index)
  
  pcts <- sweep(table(df3_long[[var_name]], df3_long$Index),
        2,
        colSums(table(df3_long[[var_name]], df3_long$Index)), '/') %>% round(3) * 100
  
  result <- matrix(paste0(counts, ' (', pcts, ')'), nrow = 2)
  
  rownames(result) <- paste0(var_name, c('_yes', '_no'))
  
  return(result)
}

area_calc <- function(var_names){
  
  first <- T
  for(area_name in var_names){
    
    print(area_name)
  

    counts <- table(df3_long[[area_name]] > 0, df3_long$Index)
    
    pcts <- sweep(table(df3_long[[area_name]] > 0, df3_long$Index),
          2,
          colSums(table(df3_long[[area_name]] > 0, df3_long$Index)), '/') %>% round(3) * 100 
    
    result <- matrix(matrix(paste0(counts, ' (', pcts, ')'), nrow = 2)[2,], nrow = 1)
    
    rownames(result) <- area_name
    
    if(first){
      first <- F
      results <- result
    } else{
      results <- rbind(results, result)
    }
  }
  
  return(results)
  
}

# cont_calc('age')
# cat_calc('race')
# 
# area_calc(c('ha_50m', 'ha_100m', 'ha_150m', 'ha_200m', 'ha_300m', 'ha_500m', 'ha_1000m'))
```

# TABLE 2

```{r}
vars <- c('age', 'race', 'gender', 'days_after_mdh', 'zha', 'i_g4salary_scaled', 'i_d9educatpar', 'currentnoflorag', 'dist_house_plant')

area <- c('ha_50m', 'ha_100m', 'ha_150m', 'ha_200m', 'ha_300m', 'ha_500m', 'ha_1000m')

NEPSY_table_2_vars <- c("nepsy", "attention_inhibitory_control", "auditory_attention_response_set", "aac", "rsc", 
                   "inhibition", "iic", "inc", "isc", 
                   "language", "cits", "snc", 
                   "memory_and_learning", "mffts", "mffdts", 
                   "visuospatial_processing", "dcgts", "gpts")

table_2_covariates <- matrix(nrow = 1, ncol = 3)
for(i in vars){
  if(i %in% c('age', 'days_after_mdh', 'zha', 'i_g4salary_scaled', 'i_d9educatpar', 'dist_house_plant')){
    table_2_covariates <- rbind(table_2_covariates, cont_calc(i))
  } else{
    table_2_covariates <- rbind(table_2_covariates, cat_calc(i))
  }
}

table_2_area <- area_calc(area)

table_2_nepsy <- matrix(nrow = 1, ncol = 3)

for(i in NEPSY_table_2_vars){
  table_2_nepsy <- rbind(table_2_nepsy, cont_calc(i))
}

table_2 <- rbind(table_2_covariates, table_2_area, table_2_nepsy)

write.csv(table_2, file = "../Figures/Final/Table_2.csv")

```

# Age 

```{r}
df3_long <- df4_long

round(tapply(df3_long$age, df3_long$Index, mean), 1)

round(tapply(df3_long$age, df3_long$Index, sd), 1)
```
# race

```{r}
table(df3_long$race, df3_long$Index)

sweep(table(df3_long$race, df3_long$Index),
      2,
      colSums(table(df3_long$race, df3_long$Index)), '/') %>% round(3)
```






# gender

```{r}
table(df3_long$gender, df3_long$Index)

sweep(table(df3_long$gender, df3_long$Index),
      2,
      colSums(table(df3_long$gender, df3_long$Index)), '/') %>% round(3)
```

# mothers day

```{r}
round(tapply(df3_long$days_after_mdh, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$days_after_mdh, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```

# ZHA 

```{r}
round(tapply(df3_long$zha, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$zha, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```
# scaled salary 

```{r}
round(tapply(df3_long$i_g4salary_scaled, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$i_g4salary_scaled, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```

# education 

```{r}
round(tapply(df3_long$i_d9educatpar, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$i_d9educatpar, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```

# cohabitation

```{r}
table(df3_long$currentnoflorag, df3_long$Index)

sweep(table(df3_long$currentnoflorag, df3_long$Index),
      2,
      colSums(table(df3_long$currentnoflorag, df3_long$Index)), '/') %>% round(3)
```

# distance

```{r}
round(tapply(df3_long$dist_house_plant, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$dist_house_plant, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```

# area within 500m

```{r}
table(df3_long$ha_500m > 0, df3_long$Index)

sweep(table(df3_long$ha_500m > 0, df3_long$Index),
      2,
      colSums(table(df3_long$ha_500m > 0, df3_long$Index)), '/') %>% round(3)
```
# area within 50m

```{r}
table(df3_long$ha_50m > 0, df3_long$Index)

sweep(table(df3_long$ha_50m > 0, df3_long$Index),
      2,
      colSums(table(df3_long$ha_50m > 0, df3_long$Index)), '/') %>% round(3)
```

# area within 100m

```{r}
table(df3_long$ha_100m > 0, df3_long$Index)

sweep(table(df3_long$ha_100m > 0, df3_long$Index),
      2,
      colSums(table(df3_long$ha_100m > 0, df3_long$Index)), '/') %>% round(3)
```
# area within 150m

```{r}
table(df3_long$ha_150m > 0, df3_long$Index)

sweep(table(df3_long$ha_150m > 0, df3_long$Index),
      2,
      colSums(table(df3_long$ha_150m > 0, df3_long$Index)), '/') %>% round(3)
```
# area within 200m

```{r}
table(df3_long$ha_200m > 0, df3_long$Index)

sweep(table(df3_long$ha_200m > 0, df3_long$Index),
      2,
      colSums(table(df3_long$ha_200m > 0, df3_long$Index)), '/') %>% round(3)
```
# area within 300m

```{r}
table(df3_long$ha_300m > 0, df3_long$Index)

sweep(table(df3_long$ha_300m > 0, df3_long$Index),
      2,
      colSums(table(df3_long$ha_300m > 0, df3_long$Index)), '/') %>% round(3)
```
# area within 750m

```{r}
table(df3_long$ha_750m > 0, df3_long$Index)

sweep(table(df3_long$ha_750m > 0, df3_long$Index),
      2,
      colSums(table(df3_long$ha_750m > 0, df3_long$Index)), '/') %>% round(3)
```
# area within 1000m

```{r}
table(df3_long$ha_1000m > 0, df3_long$Index)

sweep(table(df3_long$ha_1000m > 0, df3_long$Index),
      2,
      colSums(table(df3_long$ha_1000m > 0, df3_long$Index)), '/') %>% round(3)
```

# distance < 500

```{r}
table(df3_long$ha_500m < 500, df3_long$Index)

sweep(table(df3_long$ha_500m < 500, df3_long$Index),
      2,
      colSums(table(df3_long$ha_500m < 500, df3_long$Index)), '/') %>% round(2)
```

# attention domain 

```{r}
round(tapply(df3_long$attention_inhibitory_control, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$attention_inhibitory_control, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```

# aac 

```{r}
round(tapply(df3_long$aac, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$aac, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```
# rsc 

```{r}
round(tapply(df3_long$rsc, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$rsc, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```
# iic 

```{r}
round(tapply(df3_long$iic, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$iic, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```

# inc 

```{r}
round(tapply(df3_long$inc, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$inc, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```

# isc 

```{r}
round(tapply(df3_long$isc, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$isc, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```

# language 

```{r}
round(tapply(df3_long$language, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$language, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```

# cits 

```{r}
round(tapply(df3_long$cits, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$cits, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```

# snc 

```{r}
round(tapply(df3_long$snc, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$snc, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```
# memory/learning 

```{r}
round(tapply(df3_long$memory_and_learning, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$memory_and_learning, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```

# mffts 

```{r}
round(tapply(df3_long$mffts, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$mffts, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```

# mffdts 

```{r}
round(tapply(df3_long$mffdts, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$mffdts, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```
# visual processing 

```{r}
round(tapply(df3_long$visuospatial_processing, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$visuospatial_processing, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```

# dcgts 

```{r}
round(tapply(df3_long$dcgts, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$dcgts, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```

# gpts 

```{r}
round(tapply(df3_long$gpts, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$gpts, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```

# arts 

```{r}
# round(tapply(df3_long$arts, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)
# 
# round(tapply(df3_long$arts, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```

# overall

```{r}
vars <- c('aac', 'rsc', 'iic', 'inc', 'isc', 'cits', 'snc', 'mffts', 'mffdts', 
          'dcgts', 'gpts')

df3_long$nepsy <- rowMeans(select(df3_long, vars), na.rm = TRUE)

round(tapply(df3_long$nepsy, df3_long$Index, function(x){mean(x, na.rm = T)}), 1)

round(tapply(df3_long$nepsy, df3_long$Index, function(x){sd(x, na.rm = T)}), 1)
```
# supplemental tables - variables stratified by distance and time

```{r}
get_results <- function(value_name, distance_name, df){
  
  distance <- dplyr::pull(df, distance_name)
  value <- dplyr::pull(df, value_name)
  
  quantile_levels <- cut(distance, 
                       breaks = quantile(distance, probs = seq(0, 1, 0.25), na.rm = T), 
                       include.lowest = TRUE, 
                       labels = c("Q1", "Q2", "Q3", "Q4"))
  
  # num not na
  
  num_not_na <- sum(!is.na(distance) & !is.na(value))
  
  # means
  
  strat_means <- round(tapply(value, quantile_levels, function(x){mean(x, na.rm = T)}), 2)
  
  strat_sds <- round(tapply(value, quantile_levels, function(x){sd(x, na.rm = T)}), 2)
  
  results <- paste(strat_means, strat_sds, sep = ', ')
  
  # p-value
  

  
  the_formula <- paste(value_name, 'quantile_levels', sep = ' ~ ')
  m <- aov(formula(the_formula), data = df)
  s <- summary(m)
  p_value <- s[[1]][["Pr(>F)"]][1] %>% signif(3)
  return(c(num_not_na, results, p_value))
}

get_results_repeated <- function(value_name, distance_name, time_name, ID_name, df){
  
  df2 <- df %>% select(distance_name, value_name, time_name, ID_name) %>% na.omit()
  
  distance <- dplyr::pull(df2, distance_name)
  value <- dplyr::pull(df2, value_name)
  time <- dplyr::pull(df2, time_name)
  
  quantile_levels <- cut(distance, 
                       breaks = quantile(distance, probs = seq(0, 1, 0.25), na.rm = T), 
                       include.lowest = TRUE, 
                       labels = c("Q1", "Q2", "Q3", "Q4"))
  
  df2$quantile_levels <- quantile_levels
  
  # num not na
  
  num_not_na <- sum(!is.na(distance) & !is.na(value))
  
  # means
  
  strat_means <- round(tapply(value, list(quantile_levels, time), function(x){mean(x, na.rm = T)}), 2)
  
  strat_sds <- round(tapply(value, list(quantile_levels, time), function(x){sd(x, na.rm = T)}), 2)
  
  results <- paste(strat_means, strat_sds, sep = ', ') %>% matrix(nrow = 4, ncol = 3)
  
  colnames(results) <- c('Summer 2008', 'Spring 2016', 'Summer 2016')
  rownames(results) <- c('Q1', 'Q2', 'Q3', 'Q4')
  
  # p-value FROM REPEATED MEASURES ANOVA
  
  res.aov <- anova_test(data = df2, 
                        dv = value_name, 
                        wid = ID_name,
                        within = c(quantile_levels, time_name)
  )
  
  anova_result <- get_anova_table(res.aov)  
  
  return(list(results, anova_result))
}

dfa <- df3_long %>% filter(Index == 1)
dfb <- df3_long %>% filter(Index == 2)
dfc <- df3_long %>% filter(Index == 3)
```

```{r}
get_results_repeated('aac', 'dist_house_plant', 'Index', 'nid', df3_long)
```

# aac

```{r}
get_results('aac', 'dist_house_plant', dfa)
get_results('aac', 'dist_house_plant', dfb)
get_results('aac', 'dist_house_plant', dfc)
```


# rsc

```{r}
get_results('rsc', 'dist_house_plant', dfa)
get_results('rsc', 'dist_house_plant', dfb)
get_results('rsc', 'dist_house_plant', dfc)
```
# iic

```{r}
get_results('iic', 'dist_house_plant', dfa)
get_results('iic', 'dist_house_plant', dfb)
get_results('iic', 'dist_house_plant', dfc)
```
# inc

```{r}
get_results('inc', 'dist_house_plant', dfa)
get_results('inc', 'dist_house_plant', dfb)
get_results('inc', 'dist_house_plant', dfc)
```
# isc

```{r}
get_results('isc', 'dist_house_plant', dfa)
get_results('isc', 'dist_house_plant', dfb)
get_results('isc', 'dist_house_plant', dfc)
```
# cits

```{r}
get_results('cits', 'dist_house_plant', dfa)
# get_results('cits', 'dist_house_plant', dfb)
get_results('cits', 'dist_house_plant', dfc)
```

# snc

```{r}
get_results('snc', 'dist_house_plant', dfa)
get_results('snc', 'dist_house_plant', dfc)
```

# mffts

```{r}
get_results('mffts', 'dist_house_plant', dfa)
get_results('mffts', 'dist_house_plant', dfc)
```

# mffdts

```{r}
get_results('mffdts', 'dist_house_plant', dfa)
get_results('mffdts', 'dist_house_plant', dfc)
```

# dcgts

```{r}
get_results('dcgts', 'dist_house_plant', dfa)
get_results('dcgts', 'dist_house_plant', dfc)
```

# gpts

```{r}
get_results('gpts', 'dist_house_plant', dfa)
get_results('gpts', 'dist_house_plant', dfc)
```

